---
title: "Environments"
author: "Miguel Conde"
date: "2 de marzo de 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Intro

*Environments* formalize **relationships between variable names and values**. 

When you enter x <- 55 into the R console what you’re saying is: assign the value of 55 to a variable called x, and store this assignment in the global environment. The global environment is therefore where most R users do most of their programming and analysis.

### Creating environments with `new.env()`
```{r}
my_new_env <- new.env()
```

### Assigning pairs variables / values in an environment and retrieving them

Like a list:
```{r}
my_new_env$x <- 4
my_new_env$x
```

With `assign()` and `get()`:
```{r}
assign("y", 9, envir = my_new_env)
get("y", envir = my_new_env)
```

### Checking variables in an environment with `ls()` and `exists()`
```{r}
ls(my_new_env)
```

```{r}
exists("x", envir = my_new_env)
```

```{r}
exists("z", envir = my_new_env)
```

### Removing pairs variables / values in an environment with `rm()`
```{r}
rm(y, envir = my_new_env)
exists("y", envir = my_new_env)
```

```{r}
my_new_env$y
```

## Environments

Environments are organized in **parent/child relationships** such that every
environment keeps track of its parent, but parents are unaware of which
environments are their children.

You can see the **parents** of the global environment using the `search()` function:
```{r}
search()
```

As you can see `r search()[2]` is the parent of .GlobalEnv, and `r search()[3]` is parent of `r search()[2]`, and so on. In general the parent of .GlobalEnv is always the last package that was loaded using `library()`. 

Notice that after we load the ggplot2 package, that package becomes the parent of .GlobalEnv:

```{r}
library(ggplot2)
search()
```

### Execution Environments
Although there may be several cases where you need to create a new environment
using `new.env()`, you will more often create new environments whenever
you execute functions. An execution environment is an environment that
exists temporarily within the scope of a function that is being executed.

```{r}
x <- 10

my_func <- function(){
  x <- 5
  return(x)
}

my_func()
```
First the name `x` is bring assigned the value 10 *in the global environment*.

Then the name `my_func` is being assigned the value of the function `function(){x <- 5};return(x)}` *in the global environment*.

When `my_func()` is executed, a *new environment*, the
*execution environment*, is created **as a child** of the global environment.

This execution environment only exists while `my_func()` is running. 

Inside of the execution environment the name `x` is assigned the value `5`.

When `return()` is executed **it looks first in the execution environment** for a value that is assigned to x. Then the value 5 is returned.

In contrast to the situation above, take a look at this variation:
```{r}
x <- 10

another_func <- function(){
  return(x)
}

another_func()
```

In this situation the execution environment inside of `another_func()` does
not contain an assignment for the name `x`, **so R looks for an assignment in
the parent environment of the execution environment** which is the global
environment. Since x is assigned the value 10 in the global environment 10 is
returned.

### Complex assignment operator `<<-`

The complex assignment operator can be used to re-assign or even create name-value bindings in the parent environment (e.g., the global environment) from within a child environment (e.g., an execution environment).

```{r}
x <- 10
x
```
Re-assign:

```{r}
assign1 <- function(){
  x <<- "Wow!"
}

assign1()

x
```

Create:
```{r}
a_variable_name
```

Error in eval(expr, envir, enclos): object 'a_variable_name' not found
```{r}
exists("a_variable_name")
```

```{r}
assign2 <- function(){
  a_variable_name <<- "Magic!"
}

assign2()

exists("a_variable_name")
```

```{r}
a_variable_name
```

## Environment Access

```{r}
r_envirs <- function(ini_env) {
  print(environmentName(ini_env))
  parent_env <- parent.env(ini_env)
  if (environmentName(parent_env) != environmentName(emptyenv())) {
    r_envirs(parent_env)
  }
}

r_envirs(globalenv())
```

