---
title: "Environments"
author: "Miguel Conde"
date: "2 de marzo de 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Intro

*Environments* formalize **relationships between variable names and values**. 

When you enter x <- 55 into the R console what you’re saying is: assign the value of 55 to a variable called x, and store this assignment in the global environment. The global environment is therefore where most R users do most of their programming and analysis.

### Creating environments with `new.env()`
```{r}
my_new_env <- new.env()
```

### Assigning pairs variables / values in an environment and retrieving them

Like a list:
```{r}
my_new_env$x <- 4
my_new_env$x
```

With `assign()` and `get()`:
```{r}
assign("y", 9, envir = my_new_env)
get("y", envir = my_new_env)
```

### Checking variables in an environment with `ls()` and `exists()`
```{r}
ls(my_new_env)
```

```{r}
ls.str(my_new_env)
```


```{r}
exists("x", envir = my_new_env)
```

```{r}
exists("z", envir = my_new_env)
```

### Removing pairs variables / values in an environment with `rm()`
```{r}
rm(y, envir = my_new_env)
exists("y", envir = my_new_env)
```

```{r}
my_new_env$y
```

## Environments

Environments are organized in **parent/child relationships** such that every
environment keeps track of its parent, but parents are unaware of which
environments are their children.

You can see the **parents** of the global environment using the `search()` function:
```{r}
search()
```

As you can see `r search()[2]` is the parent of .GlobalEnv, and `r search()[3]` is parent of `r search()[2]`, and so on. In general the parent of .GlobalEnv is always the last package that was loaded using `library()`. 

Notice that after we load the ggplot2 package, that package becomes the parent of .GlobalEnv:

```{r}
library(ggplot2)
search()
```

### Execution Environments
Although there may be several cases where you need to create a new environment
using `new.env()`, you will more often create new environments whenever
you execute functions. An execution environment is an environment that
exists temporarily within the scope of a function that is being executed.

```{r}
x <- 10

my_func <- function(){
  x <- 5
  return(x)
}

my_func()
```
First the name `x` is bring assigned the value 10 *in the global environment*.

Then the name `my_func` is being assigned the value of the function `function(){x <- 5};return(x)}` *in the global environment*.

When `my_func()` is executed, a *new environment*, the
*execution environment*, is created **as a child** of the global environment.

This execution environment only exists while `my_func()` is running. 

Inside of the execution environment the name `x` is assigned the value `5`.

When `return()` is executed **it looks first in the execution environment** for a value that is assigned to x. Then the value 5 is returned.

In contrast to the situation above, take a look at this variation:
```{r}
x <- 10

another_func <- function(){
  return(x)
}

another_func()
```

In this situation the execution environment inside of `another_func()` does
not contain an assignment for the name `x`, **so R looks for an assignment in
the parent environment of the execution environment** which is the global
environment. Since x is assigned the value 10 in the global environment 10 is
returned.

### Complex assignment operator `<<-`

The complex assignment operator can be used to re-assign or even create name-value bindings in the parent environment (e.g., the global environment) from within a child environment (e.g., an execution environment).

```{r}
x <- 10
x
```
Re-assign:

```{r}
assign1 <- function(){
  x <<- "Wow!"
}

assign1()

x
```

Create:
```{r eval = FALSE}
a_variable_name
```

Error in eval(expr, envir, enclos): object 'a_variable_name' not found
```{r}
exists("a_variable_name")
```

```{r}
assign2 <- function(){
  a_variable_name <<- "Magic!"
}

assign2()

exists("a_variable_name")
```

```{r}
a_variable_name
```

## Environment in Depth
Environments can be thought of as consisting of two things:

* a *frame*, or collection of named objects, i.e., a set of symbol-value pairs.
* and an *enclosure*, a pointer to an *enclosing environment* (aka 'parent environment'). 

When R looks up the value for a symbol the frame is examined and if a matching symbol is found its value will be returned. 
If not, the enclosing environment is then accessed and the process repeated.

Environments form a tree structure in which the enclosures play the role of parents. The tree of environments is rooted in an empty environment, available through `emptyenv()`, which has no parent.

It is possible for an environment to have multiple “children”. The only environment without a parent is the empty environment.

The most common example is the frame of variables local to a function call; its **enclosure** is the environment where the function was **defined**. 

The enclosing environment or enclosure or parent environment is distinguished from the *parent frame*: the latter refers to the environment of the **caller** of a function.

```{r}
r_envirs <- function(ini_env) {
  print(environmentName(ini_env))
  parent_env <- parent.env(ini_env)
  if (environmentName(parent_env) != environmentName(emptyenv())) {
    r_envirs(parent_env)
  }
}

r_envirs(globalenv())
```

Environments are created implicitly by function calls. In this case the environment contains the variables local to the function (including the arguments), and its enclosure is the environment of the currently called function. Environments may also be created directly by new.env.

## Environment Access
When `get` or `exists` search an environment with the default `inherits = TRUE`, they look for the variable in the frame, then in the enclosing frame, and so on.

The global environment `.GlobalEnv`, more often known as the *user's workspace*, is the first item on the search path. It can also be accessed by `globalenv()`. On the search path, each item's enclosure is the next item.

The object `.BaseNamespaceEnv` is the namespace environment for the base package. The environment of the `base` package itself is available as `baseenv()`.

If one follows the chain of enclosures found by repeatedly calling `parent.env` from any environment, eventually one reaches the empty environment `emptyenv()`, into which nothing may be assigned.

The replacement function `parent.env<-` is extremely dangerous as it can be used to destructively change environments in ways that violate assumptions made by the internal C code. It may be removed in the near future.

System environments, such as the `base`, `global` and `empty` environments, have names as do the package and namespace environments and those generated by `attach()`. Other environments can be named by giving a "name" attribute, but this needs to be done with care as environments have unusual copying semantics.

* If `fun` is a function or a formula then `environment(fun)` returns the environment associated with that function or formula. If `fun` is `NULL` then the current evaluation environment is returned.

* The replacement form sets the environment of the function or formula fun to the `value` given.

* `is.environment(obj)` returns `TRUE` if and only if `obj` is an environment.

* `new.env` returns a new (empty) environment with (by default) enclosure the parent frame.

* `parent.env` returns the enclosing environment of its argument.

* `parent.env<-` sets the enclosing environment of its first argument.

* `environmentName` returns a character string, that given when the environment is printed or "" if it is not a named environment.

* `env.profile` returns a list with the following components: `size` the number of chains that can be stored in the hash table, `nchains` the number of non-empty chains in the table (as reported by HASHPRI), and `counts` an integer vector giving the length of each chain (zero for empty chains). This function is intended to assess the performance of hashed environments. When env is a non-hashed environment, `NULL` is returned.