---
title: "Expressions"
author: "Miguel Conde"
date: "2 de marzo de 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Expressions

"Encapsulated operations that can be executed in R"

An R **expression** *vector* is **a list of calls, symbols etc.**

### Code -> `quote()` -> call

Create a call from code -> `quote()`
Execute the call -> `eval()`

**Call**: *an unevaluated function call*, that is, an **unevaluated expression**

```{r}
call_2_plus_2 <- quote(2 + 2)
eval(call_2_plus_2)
```

```{r}
class(call_2_plus_2)
```

```{r}
str(call_2_plus_2)
```

```{r}
call_2_plus_2
```

### String -> `parse()` -> expression -> `deparse()` -> string

Transform a string into an expression -> `parse()`
Execute the expression -> `eval()` (also)
```{r}
str_2_plus_2 <- "2 + 2"
expr_2_plus_2 <- parse(text = str_2_plus_2)
eval(expr_2_plus_2)
```

```{r}
class(expr_2_plus_2)
```

```{r}
str(expr_2_plus_2)
```

```{r}
expr_2_plus_2
```


Transform a expression into a string -> `deparse()`
```{r}
deparse(expr_2_plus_2)
```

### Accessing and modifying calls and expressions content

```{r}
# Let's build a call object from code
sum_expr <- quote(sum(1, 5))
eval(sum_expr)
```

We can access the call content like a list:
```{r}
sum_expr[[1]]
sum_expr[[2]]
sum_expr[[3]]
```

And modify it:
```{r}
sum_expr[[1]] <- quote(paste0)
sum_expr[[2]] <- quote(4)
sum_expr[[3]] <- quote(6)
eval(sum_expr)
```

Now with a expression:
```{r}
expr_2_plus_2
```

```{r}
expr_2_plus_2[[1]]
expr_2_plus_2[[2]]
expr_2_plus_2[[3]]
```



### Composing calls with `call()`

```{r}
sum_40_50_expr <- call("sum", 40, 50)
sum_40_50_expr
```


```{r}
eval(sum_40_50_expr)
```


### Capturing calls with `match.call()`
```{r}
return_expression <- function(...){
  match.call()
  }

return_expression(2, col = "blue", FALSE)
```

### Manipulating calls

```{r}
first_arg <- function(...){
  expr <- match.call()
  
  first_arg_expr <- expr[[2]]
  first_arg <- eval(first_arg_expr)
  
  if(is.numeric(first_arg))
  {
    paste("The first argument is", first_arg)
  } 
  else 
  {
    "The first argument is not numeric."
  }
}
```

```{r}
first_arg(2, 4, "seven", FALSE)
```

```{r}
first_arg("two", 4, "seven", FALSE)
```


